---
title: 'Rising Declarative Experiment: Statical Analyses'
author: "Sunwoo Jeong"
output: html_document
---

This document provides a brief summary of the statistical results for the rising declarative data.

First, a brief note on organizing the data:

```{r}
## set working directory
setwd("/Users/sunwoojeong/Desktop/mturk-experiments/")

## Load the data
dat <- read.csv('jeong-rising-declarative-data-160828.csv')

## Summary of the data
head(dat)

## Speaker - Addressee epistemic authority on p
dat$cer_diff <- dat$sp_certainty - dat$addr_certainty

## Epistemic - Perlocutionary speaker authority
dat$auth_diff <- dat$sp_certainty - dat$sp_authority

## Subdivide data
# Only assertion/question answers for Q1
dat_2 = subset(dat, speechact_choice=="Assertion" | speechact_choice=="Question")
# Only 3rd person subject sentences
dat_target = subset(dat, sentence=="married" | sentence=="mom" | sentence=="obama" | sentence=="spaceship")
dat_target = subset(dat_target, speechact_choice=="Assertion" | speechact_choice=="Question")
# Only subject minimal pair sentences
dat_y <- dat[dat$sentence == "yemen", ]
dat_y = subset(dat_y, speechact_choice=="Assertion" | speechact_choice=="Question")

```




## Factors that influence inquisitive vs. assertive interpretations

# (column 2 of the poster)

```{r}

## Intonation and speaker gender effect

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise1", "rise2", "rise3", "qrise3", "fall"))
mylogit1 <- glm(speechact_choice ~ intonation + pronoun + speaker_gender, data = dat_target, family = "binomial")
summary(mylogit1)

## Conte(n/x)t effect

mylogit2 <- glm(speechact_choice ~ intonation + pronoun, data = dat_y, family = "binomial")
summary(mylogit2)

# Relevel conte(n/x)t factors (pronoun) to enable other direct comparisons 

dat_y$pronoun <- factor(dat_y$pronoun, levels = c("lenny", "you", "i"))
mylogit2 <- glm(speechact_choice ~ intonation + pronoun, data = dat_y, family = "binomial")
summary(mylogit2)
```

Summary of the factors that influence assertion vs. question interpretations of rising declaratives:

* Steeper rises: more question than assertion interpretations, other things being equal
* Male speakers: more assertion than question interpretations, other things being equal
* Conte(n/x)t: more assertion than question interpretations in the following order for subject personhood: I > Lenny > You




## Different conventional effects for inquisitive vs. assertive rising declaratives 

# (column 3 of the poster)

# Models on speaker epistemic certainty

Models for ratings on speaker epistemic certainty on p (Q2).

```{r}
library(lmerTest)

## Speaker epistemic certainty

dat_2$intonation <- factor(dat_2$intonation, levels = c("fall", "rise1", "rise2", "rise3", "qrise3"))
epi <- lmer(sp_certainty ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(epi)

# Relevel intonation factors to enable other direct comparisons between intonations

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise1", "rise2", "rise3", "qrise3", "fall"))
epi <- lmer(sp_certainty ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(epi)

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise2", "rise3", "qrise3", "fall", "rise1"))
epi <- lmer(sp_certainty ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(epi)

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise3", "qrise3", "fall", "rise1", "rise2"))
epi <- lmer(sp_certainty ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(epi)
```
Summary of the models on speaker epistemic certainty (on p) rating:

* Choice on speech act (Assertion vs. Question) is a significant predictor: Higher speaker epistemic authority on p when construed as an assertion
* Intonation (+ sentence type) is a significant predictor: higher epistemic certainty for falling dec > rise1 dec (>) rise2 dec, rise3 dec, rise 3 polar interrogative
* Speaker gender is not a significant predictor



# Models on speaker - addressee epistemic certainty

Models on speaker - addressee epistemic certainty on p (i.e. relative epistemic certainty for the two interlocutors; Q3-Q2) 

```{r}

## Speaker - addressee epistemic certainty

dat_2$intonation <- factor(dat_2$intonation, levels = c("fall", "rise1", "rise2", "rise3", "qrise3"))
cerdiff <- lmer(cer_diff ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(cerdiff)

# Relevel speech act factors for intercept estimate comparison (+ vs. - values)

dat_2$speechact_choice <- factor(dat_2$speechact_choice, levels = c("Question", "Assertion"))
cerdiff <- lmer(cer_diff ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(cerdiff)
dat_2$speechact_choice <- factor(dat_2$speechact_choice, levels = c("Assertion", "Question"))

# Relevel intonation factors to enable other direct comparisons between intonations

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise1", "rise2", "rise3", "qrise3", "fall"))
cerdiff <- lmer(cer_diff ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(cerdiff)

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise3", "qrise3", "fall", "rise1", "rise2"))
cerdiff <- lmer(cer_diff ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(cerdiff)

dat_2$intonation <- factor(dat_2$intonation, levels = c("qrise3", "fall", "rise1", "rise2", "rise3"))
cerdiff <- lmer(cer_diff ~ intonation + speechact_choice + speaker_gender
            + (1|speaker) + (1|workerid), data = dat_2)
summary(cerdiff)
```

Summary of the models on speaker - addressee epistemic certainty (on p) rating:

* Choice on speech act (Assertion vs. Question) is a significant predictor: speaker's epistemic authority higher than addressee's when construed as an assertion; addressee's epistemic authority higher than speaker's when construed as a question.
* Intonation (+ sentence type) is a significant predictor: higher relative speaker epistemic certainty for falling dec > rise1 dec, rise2 dec, rise3 dec, rise 3 polar interrogative
* Speaker gender is a significant predictor: male assertions with lower relative speaker epistemic certainty




# Models on speaker politeness

```{r}

## Speaker politeness

dat_2$intonation <- factor(dat_2$intonation, levels = c("fall", "rise1", "rise2", "rise3", "qrise3"))
polite <- lmer(sp_politeness ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(polite)

# Relevel intonation factors to enable other direct comparisons between intonations

dat_2$intonation <- factor(dat_2$intonation, levels = c("qrise3", "fall", "rise1", "rise2", "rise3"))
polite <- lmer(sp_politeness ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(polite)
```

Summary of the models on speaker politeness rating:

* Neither choice on speech act (Assertion vs. Question), nor speaker gender are significant predictors.
* However, sentence type + intonation is a significant predictor, particulary between falling declarative < rising declaratives < polar interrogatives: inquisitive vs. assertion rising declaratives have different politeness effects compared to their canonical alternatives (falling declarative, polar interrogative, respectively.)





# Models on speaker authority

```{r}

## Speaker authority

dat_2$intonation <- factor(dat_2$intonation, levels = c("fall", "rise1", "rise2", "rise3", "qrise3"))
authority <- lmer(sp_authority ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(authority)

# Relevel intonation factors to enable other direct comparisons between intonations

dat_2$intonation <- factor(dat_2$intonation, levels = c("qrise3", "fall", "rise1", "rise2", "rise3"))
authority <- lmer(sp_authority ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(authority)

## cf. epistemic - perlocutionary authority (Q2-Q5)

dat_2$intonation <- factor(dat_2$intonation, levels = c("rise1", "rise2", "rise3", "qrise3", "fall"))
authdiff <- lmer(auth_diff ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(authdiff)

# Relevel speech act factors for intercept estimate comparison (+ vs. - values)

dat_2$speechact_choice <- factor(dat_2$speechact_choice, levels = c("Question", "Assertion"))
authdiff <- lmer(auth_diff ~ intonation + speechact_choice + speaker_gender
                + (1|speaker) + (1|workerid), data = dat_2)
summary(authdiff)
dat_2$speechact_choice <- factor(dat_2$speechact_choice, levels = c("Assertion", "Question"))
```

Summary of the models on speaker authority rating:

* Assertion speech act associated with significantly more speaker authority than Question speech act.
* Assertive rising declaratives signal significantly less speaker authority than canonical assertive falling declaratives, whereas inquisitive rising declaratives signal comparable degree of speaker authority as canonical questioning polar interrogatives.
* Intonation a significant predictor, but speaker gender again not a significant predictor.
